---
title: "Final's project preliminary study"
output:
  html_document: 
    toc: yes
    keep_md: yes
    theme: paper
---

## Reading the libraries necessary to the project

```{r message=FALSE, warning=FALSE}
source('libs.R')
```

## Reading Power Consumption Data

```{r}
data <- read.csv('./anuario-estatistico-de-energia-eletrica/dados-anuario.csv', sep=';')
```

## Tidying data

### Converting char data to float

```{r}
data[, c('Consumidores')] <- data[, c('Consumidores')] %>% 
  str_replace_all(',(?=\\d)','.') %>% as.double
data[, c('Consumo')] <- data[, c('Consumo')] %>% 
  str_replace_all(',(?=\\d)','.') %>% as.double
```

### Converting char data to Date

```{r}
data$Data <- data$Data %>%
  as.character %>%
  strptime('%Y%m%d') %>%
  as.Date()
```




#### Grouping date

```{r}
data <- data %>% mutate(
  year = year(Data),
  year_month = yearmonth(Data),
  year_quarter = yearquarter(Data)
)
```


#### Creating a new Region Column

```{r}
data$region <- data$UF %>%
  str_replace_all('(AL|BA|PB|PE|PI|RN|SE|CE|MA)', 'NE') %>% 
  str_replace_all('(ES|SP|MG|RJ)', 'SE') %>% 
  str_replace_all('(TO|AC|AM|AP|RO|RR|PA)', 'N') %>% 
  str_replace_all('(MT|MS|DF|GO)', 'CO') %>% 
  str_replace_all('(PR|RS|SC)', 'S')
```


### Converting char data to factor

```{r}

data %>% map(class)->columns_classes
data[columns_classes[columns_classes=='character'] %>% names] <- data[columns_classes[columns_classes=='character'] %>% names] %>% as_factor()

     
# data[columns_classes[columns_classes=='character'] %>% names] %>% as_factor()
     

data %>% mutate(
  UF = as.factor(UF),
  Sistema=as.factor(Sistema),
  TipoConsumidor = as.factor(TipoConsumidor),
  SetorN1 = as.factor(SetorN1),
  SetorN2 = as.factor(SetorN2),
  TensaoN1 = as.factor(TensaoN1),
  TensaoN2 = as.factor(TensaoN2),
  TensaoN3 = as.factor(TensaoN3),
  FaixaDeConsumoN1 = as.factor(FaixaDeConsumoN1),
  FaixaDeConsumoN2 = as.factor(FaixaDeConsumoN2),) ->data



```

## Data overview

```{r}
data %>% select(-year_month, -year_quarter)  %>% summary



```


## Exploring Data

### Grouping power consumption by state

```{r}
data %>% 
  group_by(UF) %>% 
  summarise(total = sum(Consumo)) -> consumo_uf

consumo_uf
```


### Grouping consumption by month


```{r}

montlhy_consumption <- data %>% group_by(year_month) %>% 
  summarise(total=sum(Consumo))

montlhy_consumption
```



### Grouping consumption by quarter


  
```{r}
quartely_consumption <- data %>% group_by(year_quarter) %>% 
  summarize(total=sum(Consumo))
quartely_consumption
```

### Grouping consumption by year

  
```{r}
yearly_consumption <- data %>% group_by(year) %>% 
  summarize(total=sum(Consumo))
yearly_consumption
```
### Grouping by year_month and sector

```{r}
data %>% group_by(year_month,SetorN1, SetorN2 ) %>% 
  summarise(total=sum(Consumo)) %>%
  ungroup()  -> data_by_sector
```


## Reading Geographical data

```{r}
brazil_uf <- readOGR(dsn='BR', layer='BR_UF_2021')
brazil_uf@data %>% left_join(consumo_uf, by=c('SIGLA'='UF')) -> brazil_uf@data


brazil_uf %>% tm_shape() +
 tm_polygons('total',
             style='quantile',
             palette='viridis')

```


## TIME SERIES


### Time series by Sector

```{r}
data %>% 
  select(year_month, SetorN1, Consumo) %>% 
  group_by(year_month, SetorN1) %>% 
  summarise(total=sum(Consumo)) %>% 
  as_tsibble(key=SetorN1, index = year_month) -> sector_ts

sector_ts %>% autoplot() %>% ggplotly()

sector_ts %>% autoplot() 

```



```{r}
sector_ts %>% model(
  ETS = fable::ETS(),
  arima = fable::ARIMA()
) -> model_sector

model_sector %>% forecast(h='6 months') %>% 
  autoplot(sector_ts) +
  facet_grid(SetorN1 ~ .model)


model_sector %>% accuracy()
```

### Time series by Region

```{r}

data %>% 
  mutate(year_month = yearmonth(Data)) %>% 
  select(year_month, Sistema, Consumo) %>% 
  group_by(year_month, Sistema) %>% 
  summarise(total=sum(Consumo)) %>% 
  as_tsibble(key=Sistema, index = year_month)  -> region_ts

region_ts %>% autoplot() %>% ggplotly()
region_ts %>% autoplot()


```
```{r}
region_ts %>% 
  model(
    ETS = fable::ETS(),
    ARIMA = fable::ARIMA()
  ) -> model_region

model_region %>% forecast(h='6 months') %>% 
  autoplot(region_ts) +
  facet_grid(Sistema ~ .model)


model_region %>% forecast(h='16 months') %>% 
  autoplot(region_ts) +
  facet_grid(.model~Sistema)
```

  
## Visualizations

### Consumption by Consumer Type


  
```{r}
data %>%
  mutate(year = year(Data)) %>% 
  group_by(year, TipoConsumidor) %>% 
  summarise(total=sum(Consumo)) %>% 
  filter(year==2019) %>% 
  ggplot(aes(TipoConsumidor, total)) +
  geom_col(aes(fill=TipoConsumidor)) +
  coord_flip() -> consumer_type

ggplotly(consumer_type)
consumer_type
```


### Consumption by Sector N1 (Rural, Comercial, Industrial ...)

```{r}

sector_plot <- data %>% 
  mutate(year = year(Data)) %>% 
  group_by(year, SetorN1) %>% 
  summarise(total=sum(Consumo)) %>% 
  filter(year==2019) %>% 
  arrange(desc(total)) %>% 
  ggplot(aes(reorder(SetorN1, (total)), total))+
  geom_col(aes(fill=SetorN1)) +
  scale_y_continuous(labels = comma) +
  coord_flip() +
  ylab('Power Consumption (MWh)') +
  xlab('Sector') +
  labs(    title = "2019's Power Consumption by Sector", fill='Sector'  ) 


sector_plot %>% ggplotly
sector_plot 
```


### Consumption by state

```{r}

state_consumption_plot <- data %>% 
  mutate(year = year(Data)) %>% 
  group_by(year, UF) %>% 
  summarise(total=sum(Consumo)) %>% 
  filter(year==2019) %>% 
  ggplot(aes(reorder(UF, total), total))+
  geom_col(aes(fill=UF)) +
  xlab('State') +
  ylab('Total Consumption')

state_consumption_plot %>% ggplotly()

state_consumption_plot

```




### Consumption by Region (S, SE, CO, N, NE)


```{r}

region_consumption_plot <- data %>% 
  mutate(year = year(Data)) %>% 
  group_by(year, region) %>% 
  summarise(total=sum(Consumo)) %>% 
  filter(year==2019) %>% 
  ggplot(aes(reorder(region, desc(total)), total))+
  geom_col(aes(fill=region)) +
  xlab('Region') +
  ylab('Total Power Consumption') +
  labs(title = "2019's Power Consumption by Region (MWh)",
       fill = 'Region') +
  scale_y_continuous(labels = comma)

region_consumption_plot %>% ggplotly()
region_consumption_plot
```
### Annual Historic Power Consumption

```{r}
data %>% 
  group_by(year) %>% 
  summarise(total_power_consumption_by_year=sum(Consumo)) %>% 
  ggplot(aes(x=year, y=total_power_consumption_by_year)) +
  geom_line(color='blue')+
  geom_point(color='red') +
  scale_y_continuous(labels = comma) +
  labs(title = "Annual Power Total Power Consumption") +
  ylab('Total Power Consumption (MWh)') +
  xlab('Year')

```


## Other data streams

### Helper Functions

#### Apply several linear model tests
```{r}

test_linearfit <- function(model){
  
  logLik_ <- logLik(model)

  
  results <- list(
    'rsquared' = summary(model)$r.squared,
    'normality_test_p' = sf.test(model$residuals)$p.value,
    'brusch_pagan_p(>.05)' = ols_test_breusch_pagan(model)$p,
    'logLik' = logLik_
    
  )
  
  print(summary(model))
  return(results)
  
}

step_wise_k <-  3.841459

```

#### Lag several columns and creates a new dataframe

```{r}
lag_columns <- function(origin_df, columns, lag_size = 1){
  
  new_df <-  data.frame(origin_df)
  for (i in columns){
    
    new_df[,i] <- lag(new_df[,i], lag_size)
    
  }
  
  return (new_df)
    
}
```

### IPCA

### Helper functions


#### Reading IPCA


```{r}
ipca <- read.csv('http://api.bcb.gov.br/dados/serie/bcdata.sgs.433/dados?formato=csv', sep=';')

ipca %>% sample_n(10)


```


### Transforming the date input from char to Date


```{r}

ipca <- ipca %>% mutate(
  valor = valor %>% 
    str_replace(',','.') %>% 
    as.numeric(),
  data = data %>% yearmonth(format='%d-%m-%Y')
)

```


#### Grouping by periods


##### Accumulated IPCA 3 Months

```{r}

ipca['ipca_3'] <- bind_rows(
  list('ipca_3' = c(1:2)*NA),
  list('ipca_3' = rollapply((ipca$valor/100 + 1), 3, prod))) * 100 -100

```

##### Accumulated IPCA 12 Months

```{r}
ipca['ipca_12'] <- bind_rows(
  list('ipca_12' = c(1:11)*NA),
  list('ipca_12' = rollapply((ipca$valor/100 + 1), 12, prod))) * 100 -100

```


##### Accumulated IPCA 24 Months

```{r}
ipca['ipca_24'] <- bind_rows(
  list('ipca_24' = c(1:23)*NA),
  list('ipca_24' = rollapply((ipca$valor/100 + 1), 24, prod))) * 100 -100
  
```

##### Accumulated IPCA 36 Months

```{r}
ipca['ipca_36'] <- bind_rows(
  list('ipca_36' = c(1:35)*NA),
  list('ipca_36' = rollapply((ipca$valor/100 + 1), 36, prod))) * 100 -100
```


#### Filtering from Jan 2000 to Dec 2021

```{r}

ipca <- ipca %>% filter(
  data %in% c(yearmonth('2000-01-01'):yearmonth('2022-12-31'))
)
```


#### IPCA OVerview

```{r}
ipca %>% mutate(data=as.Date(data)) %>% summary

```


#### Visualizing data

##### Accumulated IPCA by different time windows

```{r}


ggplotly(
ipca %>% ggplot(aes(x=data)) +
  geom_line(aes(y=ipca_12,color='IPCA 12 months')) +
  geom_line(aes(y=ipca_3,color='IPCA 3 months')) + 
  geom_line(aes(y=ipca_24,color='IPCA 24 months')) +
  geom_line(aes(y=ipca_36,color='IPCA 36 months')) +
  xlab('Mês') +
  labs(color='IPCA', title = 'Accumulated IPCA')
)

ipca %>% ggplot(aes(x=data)) +
  geom_line(aes(y=ipca_12,color='IPCA 12 months')) +
  geom_line(aes(y=ipca_3,color='IPCA 3 months')) + 
  geom_line(aes(y=ipca_24,color='IPCA 24 months')) +
  geom_line(aes(y=ipca_36,color='IPCA 36 months')) +
  xlab('Mês') +
  labs(color='IPCA', title = 'Accumulated IPCA')
```



##### Average IPCA by different time windows


```{r}
ggplotly(
ipca %>% ggplot(aes(x=data)) +
  geom_line(aes(y=((ipca_12/100 + 1) ^(1/12))*100-100, color = 'ipca_12') ) +
  geom_line(aes(y=((ipca_3/100 + 1) ^(1/3))*100-100, color='ipca_3')) + 
  geom_line(aes(y=((ipca_24/100 + 1) ^(1/24))*100-100, color='ipca_24')) +
  geom_line(aes(y=((ipca_36/100 + 1) ^(1/36))*100-100, color='ipca_36')) +
  xlab('Month') +
  ylab('Avarage inflation') +
  labs(color='IPCA')
  
)

ipca %>% ggplot(aes(x=data)) +
  geom_line(aes(y=((ipca_12/100 + 1) ^(1/12))*100-100, color = 'ipca_12') ) +
  geom_line(aes(y=((ipca_3/100 + 1) ^(1/3))*100-100, color='ipca_3')) + 
  geom_line(aes(y=((ipca_24/100 + 1) ^(1/24))*100-100, color='ipca_24')) +
  geom_line(aes(y=((ipca_36/100 + 1) ^(1/36))*100-100, color='ipca_36')) +
  xlab('Month') +
  ylab('Avarage inflation') +
  labs(color='IPCA')


ggplotly(
montlhy_consumption %>% ggplot(aes(x=year_month))+
  geom_line(aes(y=total), color='blue', show.legend = TRUE)
)

montlhy_consumption %>% ggplot(aes(x=year_month))+
  geom_line(aes(y=total), color='blue', show.legend = TRUE)
```

#### Joining IPCA + Power Consumption datasets


```{r}
ipca_consumption <- montlhy_consumption %>% left_join(ipca, by=c(year_month="data"))
ipca_consumption
```
##### Power Consumption with lag on IPCA

- O IPCA do mês anterior, afeta o consumo de energia do mês seguinte

```{r}
lag_size = 1


ipca_consumption_lagged <- montlhy_consumption %>%
  left_join( 
    ipca %>% lag_columns(columns = c('ipca_3', 'ipca_12', 'ipca_24', 'ipca_36'), lag_size=lag_size),
    by=c(year_month = 'data')
  )

ipca_consumption_lagged


```

#### Creating linear models

##### Normal models

```{r}
lm_3 <- lm(total ~ ipca_3, ipca_consumption)
lm_12 <- lm(total ~ ipca_12, ipca_consumption)
lm_24 <- lm(total ~ ipca_24, ipca_consumption)
lm_36 <- lm(total ~ ipca_36, ipca_consumption)

```


##### Lagged models

```{r}
lm_3_lagged <- lm(total ~ ipca_3, ipca_consumption_lagged)
lm_12_lagged <- lm(total ~ ipca_12, ipca_consumption_lagged)
lm_24_lagged <- lm(total ~ ipca_24, ipca_consumption_lagged)
lm_36_lagged <- lm(total ~ ipca_36, ipca_consumption_lagged)

```

##### Grouping and evaluating models
```{r}

lm_3 %>% test_linearfit()
lm_3_lagged %>% test_linearfit()
lm_12 %>% test_linearfit()
lm_12_lagged %>% test_linearfit()
lm_24 %>% test_linearfit()
lm_24_lagged %>% test_linearfit()
lm_36 %>% test_linearfit()
lm_36_lagged %>% test_linearfit()


```
#### IPCA Impact on Residential Consumption


##### General Impact
```{r}
data_by_sector %>% filter(SetorN1=='Residencial') %>% 
  group_by(year_month) %>% 
  summarise(total= sum(total)) %>% 
  left_join(ipca, by=c(year_month='data')) -> ipca_resindential_power_consumption


lm_residential_ipca_general <- lm(total ~ valor + ipca_12 + ipca_3 + ipca_24 + ipca_36, ipca_resindential_power_consumption)


lm_residential_ipca_general %>% test_linearfit()


lm_residential_ipca_general %>% step(k=3.841459) -> lm_residential_ipca_general.s

lm_residential_ipca_general.s %>% test_linearfit()

```

##### Low income Impact
```{r}
data_by_sector %>% filter(SetorN1=='Residencial' & SetorN2=='Baixa Renda' ) %>% 
  group_by(year_month) %>% 
  summarise(total= sum(total)) %>% 
  left_join(ipca, by=c(year_month='data')) -> ipca_resindential_low_power_consumption


lm_residential_low_ipca_general <- lm(total ~ valor + ipca_12 + ipca_3 + ipca_24 + ipca_36, ipca_resindential_low_power_consumption)


lm_residential_low_ipca_general %>% test_linearfit()


lm_residential_low_ipca_general %>% step(k=3.841459) -> lm_residential_low_ipca_general.s

lm_residential_low_ipca_general.s %>% test_linearfit()

```
#### IPCA impact in all Sectors

```{r}
data_by_sector$SetorN1 %>% unique  -> SetorN1

SetorN1

results <- list()



for (i in SetorN1){
  print(i)
  data_by_sector %>% filter(SetorN1 == i) %>% 
    group_by(year_month) %>% 
    summarise(total=sum(total)) %>% 
    left_join(ipca, by=c('year_month'='data')) -> intemediary
  
  lm <- lm(total ~ valor + ipca_3 + ipca_12 + ipca_24 + ipca_36, intemediary)
  lm.s <- lm %>% step(k=3.841459)
  
  
  print(i)
  lm.s %>% test_linearfit() %>% print()
  
}
```


### Selic

API Base link: http://api.bcb.gov.br/dados/serie/bcdata.sgs.1178/dados?formato=csv

```{r}
selic_url_yearly <- 'http://api.bcb.gov.br/dados/serie/bcdata.sgs.1178/dados?formato=csv'

selic_yearly <- read.csv(selic_url_yearly, sep=';')



selic_yearly <- selic_yearly %>% 
  mutate(valor = as.numeric(str_replace(valor, ',','.')),
         data = as.Date(data, format='%d/%m/%Y'))

selic_montlhy_url <- 'http://api.bcb.gov.br/dados/serie/bcdata.sgs.4390/dados?formato=csv'


selic_montlhy <- read.csv(selic_montlhy_url, sep=';')



selic_montlhy <- selic_montlhy %>% mutate(
  valor = as.numeric(str_replace(valor, ',','.')),
         data = as.Date(data, format='%d/%m/%Y'))

```

### Selic Overview

```{r}
selic_montlhy %>% summary
```


#### Accumulating Selic

```{r}
selic_montlhy['selic_36'] <- bind_rows(
  list('selic_36' = c(1:35)*NA),
  list('selic_36' = rollapply((selic_montlhy$valor/100 + 1), 36, prod))) * 100 -100
selic_montlhy['selic_03'] <- bind_rows(
  list('selic_03' = c(1:2)*NA),
  list('selic_03' = rollapply((selic_montlhy$valor/100 + 1), 03, prod))) * 100 -100
selic_montlhy['selic_12'] <- bind_rows(
  list('selic_12' = c(1:11)*NA),
  list('selic_12' = rollapply((selic_montlhy$valor/100 + 1), 12, prod))) * 100 -100
selic_montlhy['selic_24'] <- bind_rows(
  list('selic_24' = c(1:23)*NA),
  list('selic_24' = rollapply((selic_montlhy$valor/100 + 1), 24, prod))) * 100 -100

selic_montlhy %>% relocate(selic_36, .after=selic_24) -> selic_montlhy

  selic_montlhy <- selic_montlhy %>% mutate(
    year_month= yearmonth(data)
  )


selic_montlhy %>% sample_n(15)
```

#### Sellic and Power Consumption relationship

```{r}


selic_consumption <- montlhy_consumption %>% 
  left_join(selic_montlhy) %>% 
  select(-data)
  


lm_selic_consumption <- lm(total ~ valor + selic_03 + selic_12 + selic_24 + selic_36, selic_consumption)

lm_selic_consumption %>% test_linearfit()

lm_selic_consumption.s <-lm_selic_consumption %>% step(k=3.841459)

lm_selic_consumption.s %>% test_linearfit()
```
#### Testing SELIC impact in all sectors

```{r}
data_by_sector$SetorN1 %>% unique  -> SetorN1

SetorN1

results <- list()



for (i in SetorN1){
  print(i)
  data_by_sector %>% filter(SetorN1 == i) %>% 
    group_by(year_month) %>% 
    summarise(total=sum(total)) %>% 
    left_join(selic_montlhy) %>% 
    select(-data) -> intemediary
  
  lm <- lm(total ~ valor + selic_03 + selic_12 + selic_24 + selic_36, intemediary)
  lm.s <- lm %>% step(k=3.841459)
  
  
  print(i)
  lm.s %>% test_linearfit() %>% print()
  
}



```


### PIB DATA



API Base Link: https://servicodados.ibge.gov.br/api/docs/agregados?versao=3#api-bq


#### Helper Function


##### PIB Data
```{r}

get_pib_data <- function(url){
  
  raw_data <- jsonlite::fromJSON(url)
  data <- (raw_data[[4]][[1]])[[2]][[1]]$serie
  data <- data %>% 
    pivot_longer(
      colnames(data),
      names_to = 'year',
      values_to = 'pib'
    ) %>% 
    mutate(
      pib = as.numeric(pib))
  
  return(data)
  
  
}
```





#### Reading data

##### Annual current adjusted values

```{r}

pib_url_current <- 'https://servicodados.ibge.gov.br/api/v3/agregados/6784/periodos/1996|1997|1998|1999|2000|2001|2002|2003|2004|2005|2006|2007|2008|2009|2010|2011|2012|2013|2014|2015|2016|2017|2018|2019/variaveis/9808?localidades=N1[all]'

pib_data <- get_pib_data(pib_url_current)

pib_data 
```


##### Quarterly current adjusted values

```{r}

pib_url_quarter <- 'https://servicodados.ibge.gov.br/api/v3/agregados/2072/periodos/200001|200002|200003|200004|200101|200102|200103|200104|200201|200202|200203|200204|200301|200302|200303|200304|200401|200402|200403|200404|200501|200502|200503|200504|200601|200602|200603|200604|200701|200702|200703|200704|200801|200802|200803|200804|200901|200902|200903|200904|201001|201002|201003|201004|201101|201102|201103|201104|201201|201202|201203|201204|201301|201302|201303|201304|201401|201402|201403|201404|201501|201502|201503|201504|201601|201602|201603|201604|201701|201702|201703|201704|201801|201802|201803|201804|201901|201902|201903|201904|202001|202002|202003|202004|202101|202102|202103|202104/variaveis/933?localidades=N1[all]'

quarter_pib <- get_pib_data(pib_url_quarter)

quarter_pib 


```

###### Adjusting the date 

```{r}

quarter_pib$year <- paste0(
  quarter_pib$year %>% str_match('\\d{4}'),
  '-',
  quarter_pib$year %>%
    str_match('(?<=\\d{4})\\d{2}') %>% 
    str_replace('04', '12') %>% 
    str_replace('03', '09') %>% 
    str_replace('02', '06') %>% 
    str_replace('01', '03')
  )

quarter_pib <- quarter_pib %>% 
  mutate(
    year= year %>% yearquarter()
  )

```

##### Grouping power consumption and PIB data


```{r}

lag_size_pib <- 1

pib_consumption_quarterly <- quartely_consumption %>%
  left_join(quarter_pib, by=c("year_quarter"="year"))

pib_consumption_quarterly_lagged <- quartely_consumption %>%
  left_join(quarter_pib %>% lag_columns(c('pib'),lag_size_pib )
            , by=c("year_quarter"="year"))

pib_consumption_yearly <- yearly_consumption %>%
  mutate(year = as.character(year)) %>% 
  left_join(pib_data, c(year='year'))

pib_consumption_yearly_lagged <- yearly_consumption %>%
  mutate(year = as.character(year)) %>% 
  left_join(pib_data %>% lag_columns(c('pib'), lag_size_pib),
            c('year'='year'))

pib_consumption_quarterly
pib_consumption_quarterly_lagged

```

##### Creating simple linear models for our PIB data

```{r}

pib_model_qtr <- lm(total ~ pib, pib_consumption_quarterly)
pib_model_qtr_lagged <- lm(total ~ pib, pib_consumption_quarterly_lagged)
pib_model_yrl <- lm(total ~ pib, pib_consumption_yearly)
pib_model_yrl_lagged <- lm(total ~ pib, pib_consumption_yearly_lagged)

```


##### Evaluating models

```{r}

pib_model_qtr %>% test_linearfit()
pib_model_qtr_lagged %>% test_linearfit()
pib_model_yrl %>% test_linearfit()
pib_model_yrl_lagged %>% test_linearfit()



```



##### Other links not used

- Last years annual PIB
  https://servicodados.ibge.gov.br/api/v3/agregados/6784/periodos/1996|1997|1998|1999|2000|2001|2002|2003|2004|2005|2006|2007|2008|2009|2010|2011|2012|2013|2014|2015|2016|2017|2018|2019/variaveis/9809?localidades=N1[all]
  
- Current value Per capta PIB

  https://servicodados.ibge.gov.br/api/v3/agregados/6784/periodos/1996|1997|1998|1999|2000|2001|2002|2003|2004|2005|2006|2007|2008|2009|2010|2011|2012|2013|2014|2015|2016|2017|2018|2019/variaveis/9812?localidades=N1[all]


### Reading commodities Data

- [Historical Commodities Price](https://www.indexmundi.com/commodities/)


#### Reading commodities prices

- Wheat
- Aliminum
- Coffee
- Iron Ore
- Sugar
- Crude Oil

```{r}

wheat <- read_excel('./data/commodities/wheat-240.xlsx') %>% 
  mutate(
    Price=Price %>% str_replace(',', '') %>% as.numeric()) %>% 
  mutate(
    ma_12 = rollmean(Price, 12, fill = NA, na.pad = FALSE, align = 'right'),
    ma_24 = rollmean(Price, 24, fill = NA, na.pad = FALSE, align = 'right'),
    ma_36 = rollmean(Price, 36, fill = NA, na.pad = FALSE, align = 'right'),
    Month = yearmonth(Month)
    )

aluminum <- read_excel('./data/commodities/aluminum-240.xlsx') %>% 
  mutate(
    Price=Price %>% str_replace(',', '') %>% as.numeric()  ) %>% 
  mutate(
    ma_12 = rollmean(Price, 12, fill = NA, na.pad = FALSE, align = 'right'),
    ma_24 = rollmean(Price, 24, fill = NA, na.pad = FALSE, align = 'right'),
    ma_36 = rollmean(Price, 36, fill = NA, na.pad = FALSE, align = 'right'),
    Month = yearmonth(Month)
    )

coffee <- read_excel('./data/commodities/robusta-coffee-240.xlsx') %>% 
  mutate(
    Price=Price %>% str_replace(',', '') %>% as.numeric()  ) %>% 
  mutate(
    ma_12 = rollmean(Price, 12, fill = NA, na.pad = FALSE, align = 'right'),
    ma_24 = rollmean(Price, 24, fill = NA, na.pad = FALSE, align = 'right'),
    ma_36 = rollmean(Price, 36, fill = NA, na.pad = FALSE, align = 'right'),
    Month = yearmonth(Month)
    )

iron_ore <- read_excel('./data/commodities/iron-ore.xlsx') %>% 
  mutate(
    Price=Price %>% str_replace(',', '') %>% as.numeric()) %>%
  mutate(
    ma_12 = rollmean(Price, 12, fill = NA, na.pad = FALSE, align = 'right'),
    ma_24 = rollmean(Price, 24, fill = NA, na.pad = FALSE, align = 'right'),
    ma_36 = rollmean(Price, 36, fill = NA, na.pad = FALSE, align = 'right'),
    Month = yearmonth(Month)
    )

sugar <- read_excel('./data/commodities/sugar-240.xlsx') %>% 
  mutate(
    Price=Price %>% str_replace(',', '') %>% as.numeric()) %>%
  mutate(
    ma_12 = rollmean(Price, 12, fill = NA, na.pad = FALSE, align = 'right'),
    ma_24 = rollmean(Price, 24, fill = NA, na.pad = FALSE, align = 'right'),
    ma_36 = rollmean(Price, 36, fill = NA, na.pad = FALSE, align = 'right'),
    Month = yearmonth(Month)
    )
crude_oil <- read_excel('./data/commodities/crude-oil.xlsx') %>% 
  mutate(
    Price=Price %>% str_replace(',', '') %>% as.numeric()) %>%
  mutate(
    ma_12 = rollmean(Price, 12, fill = NA, na.pad = FALSE, align = 'right'),
    ma_24 = rollmean(Price, 24, fill = NA, na.pad = FALSE, align = 'right'),
    ma_36 = rollmean(Price, 36, fill = NA, na.pad = FALSE, align = 'right'),
    Month = yearmonth(Month)
    )


```

#### Combining commodities

```{r}
wheat %>%
  select(Month, Price) %>% 
  left_join(coffee %>% select(Month, Price), by=c('Month'='Month')) %>% 
  rename(price_coffee=Price.y, price_wheat = Price.x) %>% 
  left_join(aluminum %>% select(Month, Price), by=c('Month'='Month')) %>% 
  rename(price_aluminum=Price) %>% 
  left_join(iron_ore %>% select(Month, Price), by=c('Month'='Month')) %>% 
  rename(iron_price = Price) %>% 
  left_join(sugar %>% select(Month, Price), by=c('Month'='Month')) %>% 
  rename(sugar_price = Price) %>% 
  left_join(crude_oil %>% select(Month, Price), by=c('Month'='Month')) %>% 
  rename(crude_oil = Price) %>% 
  filter(Month>=yearmonth('2007-06-01')) %>% 
  select(-Month) %>% 
  PerformanceAnalytics::chart.Correlation()





```


#### Visualizing moving averages in prices

```{r}

iron_ore %>% pivot_longer( cols = c("Price" ,"ma_12",  "ma_24",  "ma_36"))

iron_ore_plot <- iron_ore %>% select(., -Change) %>%
  pivot_longer(cols=c(-Month)) %>% 
  ggplot(aes(x=Month, y=value))+
  geom_line(aes(color=name)) +
  labs(color='Series', title= 'Iron Ore Price', y='Price BRL')

ggplotly(iron_ore_plot)
iron_ore_plot       


wheat_plot <- wheat %>% select(., -Change) %>%
  pivot_longer(cols=c(-Month)) %>% 
  ggplot(aes(x=Month, y=value))+
  geom_line(aes(color=name)) +
  labs(color='Series', title= 'Wheat Price', y='Price BRL')        

wheat_plot
wheat_plot %>% ggplotly()

coffee_plot <- coffee %>% select(., -Change) %>%
  pivot_longer(cols=c(-Month)) %>% 
  ggplot(aes(x=Month, y=value))+
  geom_line(aes(color=name)) +
  labs(color='Series', title= 'Coffee Price', y='Price BRL')


coffee_plot
ggplotly(coffee_plot)


aluminum_plot <- aluminum %>% select(., -Change) %>%
  pivot_longer(cols=c(-Month)) %>% 
  ggplot(aes(x=Month, y=value))+
  geom_line(aes(color=name)) +
  labs(color='Series', title= 'Aluminum Price', y='Price BRL')

aluminum_plot
ggplotly(aluminum_plot)


crude_oil_plot <- crude_oil %>% select(., -Change) %>%
  pivot_longer(cols=c(-Month)) %>% 
  ggplot(aes(x=Month, y=value))+
  geom_line(aes(color=name)) +
  labs(color='Series', title= 'Crude Oil Price', y='Price BRL')

crude_oil_plot
ggplotly(crude_oil_plot)

```

#### Studying the impact of coffee prices on Rural Power Consumption
```{r}
rural_consumption <- data %>% 
  filter(SetorN1=='Rural') %>% 
  group_by(year_month) %>% 
  summarise(total = sum(Consumo)) %>% 
  left_join(coffee, by=c('year_month' = 'Month'))

rural_consumption_lagged <- data %>% 
  filter(SetorN1=='Rural') %>% 
  group_by(year_month) %>% 
  summarise(total = sum(Consumo)) %>% 
  left_join(coffee %>% lag_columns(columns = c('Price', 'ma_12', 'ma_24', 'ma_36'))
            , by=c('year_month' = 'Month'))
```


##### Relationship between last month price and power consumption


```{r}
lm_rural_price_lagged <- lm(total ~ Price, rural_consumption_lagged)

lm_rural_price_lagged %>% test_linearfit()



```
##### Relationship between last 12 months coffee moving average and Power Consumption

```{r}
lm_rural_ma_12 <- lm(total ~ ma_12, rural_consumption)
lm_rural_ma_12 %>% test_linearfit()
```
##### Relationship between last 24 months coffee moving average and Power Consumption

```{r}
lm_rural_ma_24 <- lm(total ~ ma_24, rural_consumption)
lm_rural_ma_24 %>% test_linearfit()
```
##### Relationship between last 36 months coffee moving average and Power Consumption

```{r}
lm_rural_ma_36 <- lm(total ~ ma_36, rural_consumption)
lm_rural_ma_36 %>% test_linearfit()
```
##### Relationship between Price, last 12,24 and 36 months coffee moving average and Power Consumption combined


```{r}
lm_rural_total <- lm(total ~ ma_12 + ma_24 + ma_36 + Price, rural_consumption)

lm_rural_total %>% test_linearfit()

lm_rural_total.s <- lm_rural_total %>% step(k=3.841459)

lm_rural_total.s %>% test_linearfit()


lm_rural_total.s %>% olsrr::ols_vif_tol()
```

#### Rural Consumption + IPCA impact +  Coffee Price

```{r}
rural_consumption <- rural_consumption %>% left_join(ipca, by=c('year_month'='data'))


lm_rural_ipca_cofee_total = lm(total ~ . -Change, rural_consumption)

lm_rural_ipca_cofee_total %>% test_linearfit()


lm_rural_ipca_cofee_total.s <- step(lm_rural_ipca_cofee_total, k=3.841459)

lm_rural_ipca_cofee_total.s %>% test_linearfit()

lm_rural_ipca_cofee_total.s %>% olsrr::ols_vif_tol()

```
```{r}
rural_consumption_lagged <-   rural_consumption_lagged %>% 
  left_join(ipca %>% 
              lag_columns(
                columns=c('valor', 'ipca_12', 'ipca_24', 'ipca_36', 'ipca_3')),
            by=c('year_month'='data'))


lm_rural_ipca_cofee_total_lagged = lm(total ~ . -Change, rural_consumption_lagged)

lm_rural_ipca_cofee_total_lagged %>% test_linearfit()


lm_rural_ipca_cofee_total_lagged.s <- step(lm_rural_ipca_cofee_total_lagged, k=3.841459)

lm_rural_ipca_cofee_total_lagged.s %>% test_linearfit()

lm_rural_ipca_cofee_total_lagged.s %>% olsrr::ols_vif_tol()

```
#### Sugar impact in Rural Power Consumption


##### Only agroindustria
```{r}
data_by_sector %>% 
  filter(SetorN1=='Rural' & SetorN2=='Agroindústria') %>% 
  select(year_month, total) %>% 
  left_join(sugar, by=c('year_month'='Month')) -> sugar_power_consumption_ai

lm_sugar_ai <- lm(total ~ Price + ma_12+ ma_24+ ma_36, sugar_power_consumption_ai)

lm_sugar_ai %>% test_linearfit()

lm_sugar_ai %>% step(k=3.841459) -> lm_sugar_ai.s


lm_sugar_ai.s %>% test_linearfit()
```

##### All Rural sector
```{r}
data_by_sector %>% 
  filter(SetorN1=='Rural') %>% 
  group_by(year_month) %>% 
  summarise(total=sum(total)) %>% 
  select(year_month, total) %>% 
  left_join(sugar, by=c('year_month'='Month')) -> sugar_power_consumption

lm_sugar <- lm(total ~ Price + ma_12+ ma_24+ ma_36, sugar_power_consumption)

lm_sugar %>% test_linearfit()

lm_sugar %>% step(k=3.841459) -> lm_sugar.s


lm_sugar.s %>% test_linearfit()
```



#### Crude Oil Impact in Power Consumption

```{r}


data_by_sector %>% group_by(year_month) %>% 
  summarise(total=sum(total)) %>% 
  left_join(crude_oil, by=c('year_month'='Month')) -> power_consumption_crude_oil


power_consumption_crude_oil %>% select(-year_month, -Change) %>% chart.Correlation()

lm_oil <- lm(total ~ Price + ma_12+ ma_24+ ma_36, power_consumption_crude_oil)


lm_oil %>% step(lm_oil, k = 3.841459) -> lm_oil.s


lm_oil.s %>% test_linearfit()
```



#### Studying the impact of inflation in commercial power consumption

```{r}
comercial_consumption <- data_by_sector %>%
  filter(SetorN2 == 'Comercial N2') %>% 
  select(year_month, total) %>%
  left_join(ipca, c(year_month='data'))

```

##### Testing linear fit for all time windows

```{r}
lm_comercial_total <- lm(total ~ ., comercial_consumption)
lm_comercial_total %>% test_linearfit()
```

###### Stepwise method

```{r}
lm_comercial_total.s <- lm_comercial_total %>% step(k = 3.841459)
lm_comercial_total.s %>% test_linearfit()
```


#### Industrial and iron-ore relationship

```{r}
data_by_sector %>% filter(SetorN1 == 'Industrial') %>% select(-SetorN2) %>% 
  left_join(iron_ore, by=c(year_month='Month')) -> industrial_iron_price

industrial_iron_price

industrial_model <- lm(total ~ Price +ma_12 +ma_24 +ma_36 , industrial_iron_price)

industrial_model %>% test_linearfit()

industrial_model %>%  step(k = 3.841459) -> industrial_model.s

industrial_model.s %>% test_linearfit()

```

## Reverse impact

```{r}
data_by_sector %>% group_by(year_month,SetorN1) %>% 
  summarise(total=sum(total))->data_by_sector_N1


data_by_sector_N1  %>% group_by(SetorN1) %>%
  mutate(setor_avg_03 = rollmean(total, 03, fill= NA, align = 'right'),
        setor_avg_06 = rollmean(total, 06, fill= NA, align = 'right'),
        setor_avg_12 = rollmean(total, 12, fill= NA, align = 'right'),
        setor_avg_24 = rollmean(total, 24, fill= NA, align = 'right'),
        setor_avg_36 = rollmean(total, 36, fill= NA, align = 'right'), ) %>%
  arrange(SetorN1, year_month) -> data_by_sector_N1



```

### Visualizations

```{r}


comercial_plot <-  data_by_sector_N1 %>% 
            ungroup() %>% 
            filter(SetorN1 == 'Comercial') %>% 
            pivot_longer(cols = data_by_sector_N1[,-c(1:2)] %>% names) %>% 
            ggplot(aes(x=year_month, y=value))+
            geom_line(aes(color=name)) +
            labs(title='Comercial')

comercial_plot
ggplotly(comercial_plot)


consumo_proprio_plot <-    data_by_sector_N1 %>% 
            ungroup() %>% 
            filter(SetorN1 == 'Consumo Próprio') %>% 
            pivot_longer(cols = data_by_sector_N1[,-c(1:2)] %>% names) %>% 
            ggplot(aes(x=year_month, y=value))+
            geom_line(aes(color=name)) +
            labs(title='Consumo Próprio')
consumo_proprio_plot
ggplotly(consumo_proprio_plot)


iluminacao_publica_plot <-    data_by_sector_N1 %>% 
            ungroup() %>% 
            filter(SetorN1 == 'Iluminação Pública') %>% 
            pivot_longer(cols = data_by_sector_N1[,-c(1:2)] %>% names) %>% 
            ggplot(aes(x=year_month, y=value))+
            geom_line(aes(color=name)) +
            labs(title='Iluminação Pública')
iluminacao_publica_plot

ggplotly(iluminacao_publica_plot)

industrial_plot <- data_by_sector_N1 %>% 
            ungroup() %>% 
            filter(SetorN1 == 'Industrial') %>% 
            pivot_longer(cols = data_by_sector_N1[,-c(1:2)] %>% names) %>% 
            ggplot(aes(x=year_month, y=value))+
            geom_line(aes(color=name)) +
            labs(title='Industrial')
industrial_plot
ggplotly(industrial_plot)

poder_publico_plot <-    data_by_sector_N1 %>% 
            ungroup() %>% 
            filter(SetorN1 == 'Poder Público') %>% 
            pivot_longer(cols = data_by_sector_N1[,-c(1:2)] %>% names) %>% 
            ggplot(aes(x=year_month, y=value))+
            geom_line(aes(color=name)) +
            labs(title='Poder Público')
poder_publico_plot
ggplotly(poder_publico_plot)

residencial_plot <-    data_by_sector_N1 %>% 
            ungroup() %>% 
            filter(SetorN1 == 'Residencial') %>% 
            pivot_longer(cols = data_by_sector_N1[,-c(1:2)] %>% names) %>% 
            ggplot(aes(x=year_month, y=value))+
            geom_line(aes(color=name)) +
            labs(title='Residencial')

residencial_plot
ggplotly(residencial_plot)

rural_plot <-    data_by_sector_N1 %>% 
            ungroup() %>% 
            filter(SetorN1 == 'Rural') %>% 
            pivot_longer(cols = data_by_sector_N1[,-c(1:2)] %>% names) %>% 
            ggplot(aes(x=year_month, y=value))+
            geom_line(aes(color=name)) +
            labs(title='Rural')

rural_plot
ggplotly(rural_plot)


servico_publico_plot <-   data_by_sector_N1 %>% 
            ungroup() %>% 
            filter(SetorN1 == 'Serviço Público') %>% 
            pivot_longer(cols = data_by_sector_N1[,-c(1:2)] %>% names) %>% 
            ggplot(aes(x=year_month, y=value))+
            geom_line(aes(color=name)) +
            labs(title='Serviço Público')
servico_publico_plot
ggplotly(servico_publico_plot)


```

### Impact on IPCA

```{r}
data_by_sector_N1 %>%
  group_by(year_month, SetorN1) %>%
  select(setor_avg_12, total) %>%
  ungroup() %>%
  pivot_wider(names_from = SetorN1, 
              values_from = c(total,setor_avg_12)) %>% 
  drop_na() -> data_by_sector_N1.avg12
```

```{r}
data_by_sector_N1.avg12 <- data_by_sector_N1.avg12 %>% left_join(ipca_consumption %>% select(year_month, valor))


lm_consumption_ipca <- lm(valor ~ . -year_month, data_by_sector_N1.avg12)

lm_consumption_ipca %>% test_linearfit()

lm_consumption_ipca <- lm_consumption_ipca %>% step(k=step_wise_k)

lm_consumption_ipca %>% test_linearfit()

```
### Impact on PIB

#### Grouping by quarter


```{r}
data_by_sector_N1 %>% mutate(
  quarter = yearquarter(year_month)
) %>% 
  group_by(quarter, SetorN1) %>% 
  summarise(total=mean(total)) %>%
  ungroup() %>% 
  pivot_wider(names_from = 'SetorN1', values_from = total) -> data_by_sector_N1_quarter
```

#### Lagging values

```{r}
lag_cols_qtr <- c("Comercial","Consumo.Próprio","Iluminação.Pública","Industrial","Poder.Público","Residencial","Rural","Serviço.Público")

data_by_sector_N1_quarter.l <- data.frame(data_by_sector_N1_quarter)




data_by_sector_N1_quarter.l[,"Comercial"] <- data_by_sector_N1_quarter.l[,"Comercial"] %>% lag(1)
data_by_sector_N1_quarter.l[,'Consumo.Próprio'] <- data_by_sector_N1_quarter.l[,'Consumo.Próprio'] %>% lag(1)
data_by_sector_N1_quarter.l[,"Iluminação.Pública"] <- data_by_sector_N1_quarter.l[,"Iluminação.Pública"] %>% lag(1)
data_by_sector_N1_quarter.l[,"Industrial"] <- data_by_sector_N1_quarter.l[,"Industrial"] %>% lag(1)
data_by_sector_N1_quarter.l[,"Poder.Público"] <- data_by_sector_N1_quarter.l[,"Poder.Público"] %>% lag(1)
data_by_sector_N1_quarter.l[,"Residencial"] <- data_by_sector_N1_quarter.l[,"Residencial"] %>% lag(1)
data_by_sector_N1_quarter.l[,"Rural"] <- data_by_sector_N1_quarter.l[,"Rural"] %>% lag(1)
data_by_sector_N1_quarter.l[,"Serviço.Público"] <- data_by_sector_N1_quarter.l[,"Serviço.Público"] %>% lag(1)

data_by_sector_N1_quarter.l %>%
  left_join(quarter_pib, c('quarter'='year')) %>% 
  drop_na() -> consumption_pib_quarter


consumption_pib_quarter
```

#### Testing impact on PIB

```{r}
lm_consumption_pib <- lm(pib ~ . -quarter, consumption_pib_quarter)

lm_consumption_pib %>% test_linearfit()

lm_consumption_pib.s <- lm_consumption_pib %>% step(k=step_wise_k)


lm_consumption_pib.s %>% test_linearfit()

```

